%This is the acknowledgments tex file, which should have been set in the main file using the command \verb|\setAcknowledgementsFile{Drive:/file/location/acknowledgementsFile}|.

%Keep in mind this should be written in first person, eg;

%This template owes thanks to all the LaTeX Specialists who have helped contribute over the years. A big thanks to Ron Smith, James Booth, Jason Nowell, December Slater, and Meg Renard for their contributions and maintenance of this document.  

% \begin{itemize}
%     \item Family -- Dad, Mom, Jack, Hannah, Nonni and Boppi, Kelly and Brian; plus extended -- mention upbringing in Burlington, CT
%     \item UConn mentors -- Joo, Krawec, Fuller, Herzberg -- others?
%     \item Tom 
%     \item Committee -- VB, SR, PT, JB
%     \item FICS lab -- KB, students past and present
%     \item ETHZ Applied Cryptography group: Kenny, Mia, Nico -- more by name?
%     \item TUD Cryptoplexity Group  -- Marc and Moritz
%     \item Friends -- Gainesville (Tim, JT, Jefferson, Logan, Gavin, Trivia Group), CT (all in Rod Stewart Chat), high school, all those I forgot to name. 
%     \item Julia
% \end{itemize}

This is the last section of my dissertation, and somehow, it also feels like the toughest (I would be remiss to forget anyone!). While defending a dissertation and earning a doctorate may seem like a solitary achievement, I can unequivocally assure you that it is not. This accomplishment (and my life trajectory in general) would not have been possible without a wonderful and unwavering support system, both personal and professional.

I would like to begin by thanking my parents, Dawn and Tim. They have been nothing but loving and encouraging since the day I was born and gave me an almost idyllic childhood in the wonderful small town of Burlington, Connecticut. Instilling in me and my siblings, Jack and Hannah (whom I also owe a lot to), the values of education and hard work was a gift beyond measure. I also want to thank the entirety of my (very large) extended family -- you have forever impressed upon me the importance of family and gathering. I especially want to acknowledge my maternal grandparents, my Nonni and Boppi. They are my only local relatives in north-central Florida, and I have greatly enjoyed the many dinners and visits over the past five years. It is with great sadness that my Boppi is not here to witness the end of my doctoral journey, but his memory is a blessing, and I am comforted by the thought that he is at peace. Lastly, to my soon-to-be in-laws, Kelly and Brian: thank you for all of your support over the last five years (especially the plane tickets to visit home), for making me a part of your family, and above all, for giving me the greatest gift of all -- my fiancée Julia.

Professionally, I have had many mentors who encouraged my intellectual curiosity and guided my growth as a scientist. At the University of Connecticut, I am grateful to Dr. Kyungseon Joo for introducing me to research and offering incredible opportunities to a wide-eyed underclassman. Dr. Walter Krawec welcomed me to his quantum key distribution lab, redirecting me from physics toward computer science and cybersecurity. I also thank him for his mentorship on my first academic papers. Lastly, I want to thank Dr. Benjamin Fuller and Dr. Amir Herzberg for introducing me to cryptography and inspiring my passion for it.

I would not be who I am today without Dr. Thomas Shrimpton. Quite literally, meeting Tom at Real World Crypto 2020 convinced me to go to the University of Florida and work with him. I am deeply thankful for his guidance over the past five years and for suggesting (or perhaps imposing upon me) the research topic that forms this dissertation. I will never forget our sessions at the chalkboard, marathon Slack chats, and all-night paper pushes. Although Tom is not my advisor of record due to his bewildering decision to move to industry (I jest, of course), he deserves the most credit for my development as a researcher.

I also extend my thanks to my committee -- Dr. Vincent Bindschaedler, Dr. Patrick Traynor, Dr. Sara Rampazzi, and Dr. Jeremy Booher -- for their encouragement and helpful suggestions along the way. I am especially grateful to Vincent for stepping in as my advisor and helping (or dragging) me across the finish line. I also want to recognize Dr. Kevin Butler, who, although not on my committee, provided invaluable leadership at the Florida Institute for Cybersecurity (FICS). To all the faculty, administrators, and students (past and present) of FICS, as well as to the institute itself -- my deepest gratitude. It is a truly wonderful place to work, and I could not have asked for a better lab to complete my doctorate in. While the volume and impact of research coming out of FICS is impressive, it is the people and the culture that make it exceptional. You will not find a group of smarter, more innovative, friendly, and honest individuals. I am honored to have been a small part of it.

During my PhD, I was fortunate to participate in two international collaborations. I am especially thankful to Dr. Kenneth Paterson (ETH Zurich, Applied Cryptography Group) and Dr. Marc Fischlin (TU Darmstadt, Cryptoplexity Group) for hosting me for research visits and collaborating on works that appear in this dissertation. Mia Filić (ETH Zurich) deserves special recognition -- she is my longest collaborator and a dear friend. Without her, this work would not have been possible. I also wish to acknowledge Moritz Huppert (TU Darmstadt); he has been an insightful and tenacious collaborator and deserves credit for introducing me to many delicious German beers (and his more limited success at convincing me of the joys of German humor). I also thank all the students of these groups -- they were incredibly welcoming during my visits, and I am glad to call many of you friends. A special thanks to Nicholas-Philip Brandt (ETH Zurich); although Nico and I did not collaborate on work in this dissertation, he is one of the most meticulous researchers I know.

Beyond my family and professional mentors, this journey would not have been possible without a wide and geographically dispersed group of friends. To Tim, Jefferson, Logan, Gavin, and John -- thank you for making Gainesville fun these past five years. To my UConn friends Jake, Andy, Owen, Mitchell, Kevin, Skippy, Stofko, Bucco, and Paul -- thank you for all the shenanigans, for visiting me in Florida, and for Huskies basketball. To my childhood friends Cameron, Jack, and Alex -- thank you for making it feel like no time has passed when we get together. To everyone who has ever played a round of bar trivia with me, and to all my other dear friends whom I could not possibly enumerate -- thank you.

Lastly, and above all, thank you to my Julia. Your partnership over the past decade has kept me steady. Thank you for your unwavering support (even when I decided to move over a thousand miles away to pursue a PhD with some guy named Tom I met at a conference). You are everything I could have dreamed of in a partner. I am beyond excited to marry you and build our life together. I love you with all my heart.